On the Dynamics of Anthropomorphism in Robotics

### 2013-07-12 dynamics-anthropomorphism.tex
changes

Literature on anthropomorphism in robotics is quite diverse and one strives hard to extract a coherent conclusion. One reason for this might be that robotics is a multi-disciplinary field and researchers from very different domains might integrate diverse or even contradictory understandings of the phenomenon. Despite the fact that there is no commonly accepted definition of anthropomorphism in robotics, and also the terminology is not clear, the terms \textit{anthropomorphic} or \textit{human-like} are often used as if their meanings were clear and agreed upon \cite{persson_anthropomorphism_2000}. Duffy \cite{duffy_anthropomorphism_2002} even argues that the terms might be misused. For instance, some researchers refer to \textit{"the robot's level of anthropomorphism"} \cite{bartneck_is_2007}, whereas others argue that the system or artifact itself does not "contain anthropomorphism" \textit{per se} but only gives rise to the process of anthropomorphizing in a given user and situation. Consequently, anthropomorphism emerges in the \textit{interaction} between the technology and the user \cite{persson_anthropomorphism_2000}.

Paradoxically, anthropomorphism in the perception of non-human agents (such as robots) is commonly observed but at the same time poorly understood \cite{epley_seeing_2007}. Why? The phenomenon itself has been found to be very complex. Still, it can sometimes also be subtle and as such hard to study. What might also account for the rather poor (or controversial) understanding of anthropomorphism in robotics, is the limited reliability of some findings (or interpretations) from experiments investigating anthropomorphism. Most of these experiments are conducted in controlled lab-settings during short-term human-robot interactions. This experimental setting (short term and laboratory context) can be critical when studying a social phenomenon like anthropomorphism and possibly lead to over-interpretations. Some studies also seem to contrast two or more different anthropomorphic and mechanical systems that might not be appropriate for comparison under the given context and research question. \hl{say more}

To sum it up, while anthropomorphism in robotics is a commonly discussed and studied trait of human-robot interaction, we think it is paradoxically an overlooked research topic. The understanding of anthropomorphism in robotics seems to not take into account the wide range of phenomena that it encompasses. The traditional understanding of anthropomorphism in robotics to date considers two main factors that account for the social phenomenon: the characteristics of \textbf{1) the robot's design} (degree of human-likeness) and the psychological determinants in \textbf{2) the human user} (see \cite{epley_seeing_2007}). It has been shown that both these factors can facilitate or hinder anthropomorphism. We propose to go beyond this traditional perception of anthropomorphism as a static feature of HRI: Based on Persson \textit{et al.}'s~\cite{persson_anthropomorphism_2000} argumentation that anthropomorphism is a multi-layered phenomenon which arises in different levels, we suggest to understand anthropomorphism as a \emph{dynamic}, \emph{context-dependent} process. To this end, we apply Persson \textit{et al.}'s six levels on anthropomorphism to HRI and introduce three \emph{interaction phases}. These interaction phases of anthropomorphism reflect also the results of a long-term study that we conducted in a real human environment and provide an account of the so-called \textit{novelty effect}. We also draw attention to two new determinants of anthropomorphism, namely, the context of use and the fact of getting used to a system over time. We propose to add these two aspects as a third factor, the characteristics of \textbf{3) the interaction} (time and space), to the previous two factors of anthropomorphism (robot and user).

>>>>

Literature on anthropomorphism (the social phenomenon) in robotics as well as on anthropomorphic design of robots is quite diverse and one strives hard to extract a coherent conclusion. One reason for this might be that robotics is a multi-disciplinary field and researchers from very different domains might integrate diverse or even contradictory understandings, or simply refer to different things. However, despite the fact that there is no commonly accepted definition, and also the terminology is not clear, the terms \textit{anthropomorphism}, \textit{anthropomorphic} or \textit{human-like} are often used as if their meanings were clear and agreed upon \cite{persson_anthropomorphism_2000}. Duffy \cite{duffy_anthropomorphism_2002} argues that these terms might even be misused (in accordance with \cite{epley_when_2008}). For instance, some researchers refer to \textit{"the robot's level of anthropomorphism"} \cite{bartneck_is_2007}, whereas others argue that the system or artifact itself does not "contain anthropomorphism" \textit{per se} but only gives rise to the process of anthropomorphizing in a given user and situation. Consequently, Persson \textit{et al.} conclude that anthropomorphism emerges in the \textit{interaction} between the technology and the user \cite{persson_anthropomorphism_2000}.
\hl{[In this article, we accept this latter viewpoint and would like to clearly distinguish between the term \textit{anthropomorphism} as a social phenomenon and \textit{anthropomorphic} as a specific design. This however, is not yet a definition of what exactly both terms refer to. Still, we acknowledge that both terms can refer to different levels or types; which we will outline later.]}

Paradoxically, human's tendency to anthropomorphize non-human agents (such as robots) is commonly observed but at the same time it is rather poorly understood \cite{epley_seeing_2007}. There might be several reasons. First, probably because the phenomenon itself has been found to be very complex though sometimes subtle and as such hard to study. Second, the rather poor (or controversial) understanding, might also be due to the limited reliability of some findings (or interpretations) from experiments investigating anthropomorphism. Most of the experiments that investigate anthropomorphism in robotics are conducted in controlled lab-settings during short-term human-robot interactions. This experimental setting can be critical when studying a social phenomenon like anthropomorphism and possibly lead to over-interpretations. Some studies (e.g. exploring characteristics of the "uncanny valley") also seem to contrast two or more different anthropomorphic or mechanical systems that might not be appropriate for comparison given the complex context. Others might also use empirical methods inappropriate for the given research question which leads to inaccurate results. \hl{say more}

To sum it up, while anthropomorphism in robotics is a commonly discussed and studied trait of human-robot interaction, we think it paradoxically lacks a formal ground and unified understanding. The understanding of anthropomorphism in robotics seems to not take into account the wide range of phenomena that it encompasses. The traditional understanding of anthropomorphism in robotics to date considers two main factors that account for the social phenomenon: the characteristics of \textbf{1) the robot's design} (degree of human-likeness) and the psychological determinants in \textbf{2) the human user} (see \cite{epley_seeing_2007}). It has been shown that both these factors can facilitate or hinder anthropomorphism. We propose to go beyond this traditional perception of anthropomorphism as a static feature of HRI: Based on Persson \textit{et al.}'s~\cite{persson_anthropomorphism_2000} argumentation that anthropomorphism is a multi-layered phenomenon which arises in different levels, we suggest to understand anthropomorphism as a \emph{dynamic}, \emph{context-dependent} process. To this end, we apply Persson \textit{et al.}'s six levels on anthropomorphism to HRI and introduce three \emph{interaction phases}. These interaction phases of anthropomorphism reflect also the results of a long-term study that we conducted in a real human environment and provide an account of the so-called \textit{novelty effect}. We also draw attention to two new determinants of anthropomorphism, namely, the context of use and the fact of getting used to a system over time. We propose to add these two aspects as a third factor, the characteristics of \textbf{3) the interaction} (time and space), to the previous two factors of anthropomorphism (robot and user).
